Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-20 doc/report/msc-report.tex
doc/report/msc-report.tex: * cosmetic changes. * general shaping. * added Storage / Data Model section to Implementation chapter. needs work. * additions to Introduction chapter.
doc/report/references.bib:
* changed OLAP data scalability to book, for looks.
* added Omniture.

Making sense of data at a massive scale is one of the biggest challenges we face
at Last.fm. Interpreting patterns in user behaviour becomes a challenge when
millions of users interact in billions of combinations; the data sets must be
analyzed, summarized and presented visually in order to convey their meaning.

@@

Inevitably, the services and operations will generate a log for later
analysis. As the services grow in size, so will the data sets they generate.

@@

The size of the data sets at last.fm is in the range of several gigabytes per
day. Some data sets are many times larger. It is our belief that this is
typical.

@@

The International Aerodyne Corporation, a leading exporter of woodworking
nails, uses a spreadsheet to keep track of the number of nails exported to
each country every month (see figure).

@@

Last.fm is an internet radio station focusing on content, being a musical
encyclopedia. The web site.

Last.fm lets users submit what music they listen to, called \textit{scrobbling}.

Radio streams to users.

@@

OLAP is something. It's drawbacks are unknown.

@@

Omniture is used at Last.fm etc.

@@

% TODO: make correct.
The following overview is entirely incorrect:

Chapter 2 establish the theoretical base of the problem statement.

In chapter 3, the tools used for the implementation are declared, moving
over to chapter 4 where the details of the implementation are presented.

How to use the Zohmg system is presented in chapter~5, while chapter 6
shows an example work flow.

Chapters 7 and 8 discuss the results, related, and future work, and
presents the conclusion, respectively.

@@

Secondly, this chapter features new techniques in distributed storage and
computing; these techniques are Google's BigTable and MapReduce.

@@

OLAP\footnote{On-Line Analytical Processing} is an umbrella term for systems
built in a certain manner and that are optimized quickly answer
multi-dimensional analytical queries.

@@

% TODO: fix positioning.
\begin{table}[h]
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \small \bf Row Key & \small \bf Time & \small \bf Column &
        \multicolumn{2}{|c|}{\small \bf Column} \\
         & \small \bf Stamp & \small "contents:" &
        \multicolumn{2}{|c|}{\small "anchor:"} \\
        \hline
         & t9 & & \small "anchor:cnnsi.com" & "CNN" \\
         & t8 & & \small "anchor:my.lok.ca" & "CNN.com" \\
        "com.cnn.www" & t6 & \small "{\textless}html\textgreater..." & & \\
         & t5 & \small "{\textless}html\textgreater..." & & \\
         & t3 & \small "{\textless}html\textgreater..." & & \\
        \hline
        \end{tabular}
        \caption{Example HBase table.}
    \end{center}
\end{table}

@@

There exists a few databases which inspired by BigTable; they aim to be
distributed and have column-oriented properties: HBase, Cassandra, MonetDB,
MongoDB, Hypertable. CouchDB is an document-oriented database.

@@

\begin{lstlisting}[caption=Example log line entry for the Apache HTTP Server's
Combined Log Format.,captionpos=b]
   127.0.0.1 - emma [20/Apr/2009:13:37:00 +0100]
   "GET /music HTTP/1.0" 200 2326 "http://last.fm/"
   "Mozilla/5.0 (X11; U; Linux i686)"
\end{lstlisting}

@@

\begin{lstlisting}[language=Python,caption=An example map function.,captionpos=b]
   def map(key, value):
      log = parse_apache_log(value)
      dimensions = {'useragent' : log.useragent,
                    'path'      : log.path,
                    'status'    : log.statuscode}
      values = {'pageviews' : 1,
                'bytes'     : log.size}

@@

Since the wish is to store projections rather than all combinations of all
dimensions, the data needs to be mangled slightly before sending it down
the pipe to the underlying data store. Dumbo expects a key-value pair from the
mapper. Hadoop will collect all keys and give the reducer a key and an iterator
of values.

@@

Remember, the usermapper gives us a point in n-space and a list of
measurements. We need to break out the measurements and sum over each such
measurement. Zohmg solves this by wrapping the user map function.

@@

@@

The data export server is built with a WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
framework, which includes both middleware and application parts. The middleware
applications are responsible for dispatching requests to the corresponding
application or file. WSGI middleware receives a WSGI request and then performs
logic upon this request, before passing the request to a WSGI application.
\cite{definitive_guide_to_pylons,paste} Since the main goal was not to build a
middleware framework from scratch, an existing such was used.

@@

The data export server is built with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URL it was fed. \cite{python_paste}

@@

\textbf{Serializing} &
Serializing data is the task of converting an object into a sequence of bits,
making it possible to store on a storage medium or transmitted over a network
link. Rereading the sequence of bits should restore the original state of the
object it was in before the serialization. \\
\end{tabular}

@@

\section{Typical use case}

The typical use case for Zohmg is to read one or more column-qualifiers,
which correspond to one or more points in n-space, over a range of time.
Therefor, the keys are made up of the time and unit of the measurement,
and the qualifier queried for represents the point in n-space. More on
this in the data model section.

>>>>

Analyzing data at a massive scale is one of the biggest challenges we face at
Last.fm. Interpreting patterns in user behaviour becomes a challenge when
millions of users interact in billions of combinations; the data sets must be
analyzed, summarized and presented visually in order to convey their meaning.

@@

Generally, the services and operations will generate a log for later analysis.
As the services grow in size, so will the data sets they generate.

@@

The size of the data sets at Last.fm is in the range of several gigabytes per
day. Some data sets are many times larger. It is our belief that this is
typical.

@@

The International Aerodyne Corporation, a leading exporter of woodworking
nails, uses a spreadsheet to keep track of the number of nails exported to
each country every month (see table 1.1).

@@

Last.fm is an internet radio station focusing on content, being a musical
encyclopedia. Last.fm lets users submit what music they listen to, called
\textit{scrobbling}, this user submitted data is analyzed and Last.fm offer
their users personalized charts, music and event recommendations, based ther
listening. Last.fm also streams content, via internet, to users. The Last.fm web
site ties together all these areas and presents it to the users.

@@

OLAP\footnote{On-Line Analytical Processing} is a broad category of techniques
for processing data, and visualizing the analysis. It is called online because
the data store will be queried, and the user user expects a, at least close to,
direct response.

@@

Omniture is a data warehouse used at Last.fm. Criticism against Omniture is
that it is cumbersome and slow to use. \cite{omniture}

@@

In chapter 2 general data warehous concepts are introduced; it also introduces
new techniques that the reader might not know.

Chapter 3 establish the theoretical base of the problem statement, and the
expected system's function and requirements.

Chapter 4 presents the work method during the thesis. It also presents tools
possible to use for solving the problem statement.

In chapter 5, the implementation choices and details of the Zohmg system are
presented.

Chapters 6 discuss the results, related, future work, and presents the
conclusion.

In appendix A an example work flow is presented. Appendix B contains a glossary.

@@

The last section features new techniques in distributed storage and computing;
these techniques are Google's BigTable and MapReduce.

@@

OLAP is an umbrella term for systems built in a certain manner and that are
optimized quickly answer multi-dimensional analytical queries.

@@

% TODO: fix positioning.
\begin{table}[h]
    \begin{center}
        \begin{tabular}{|l|l|l|l|l|}
        \hline
        \small \bf Row Key & \small \bf Time & \small \bf Column &
        \multicolumn{2}{|c|}{\small \bf Column} \\
         & \small \bf Stamp & \small "contents:" &
        \multicolumn{2}{|c|}{\small "anchor:"} \\
        \hline
         & t9 & & \small "anchor:cnnsi.com" & "CNN" \\
         & t8 & & \small "anchor:my.look.ca" & "CNN.com" \\
        "com.cnn.www" & t6 & \small "{\textless}html\textgreater..." & & \\
         & t5 & \small "{\textless}html\textgreater..." & & \\
         & t3 & \small "{\textless}html\textgreater..." & & \\
        \hline
        \end{tabular}
        \caption{Example HBase table.}
    \end{center}
\end{table}

@@

There exists a few databases which inspired by BigTable; they aim to be
distributed and have column-oriented properties: HBase, Cassandra, MonetDB,
MongoDB, Hypertable. CouchDB is a document-oriented database.

@@

\begin{lstlisting}[caption=Example log line entry for the Apache HTTP Server's
Combined Log Format.,captionpos=b]
   127.0.0.1 - emma [20/Apr/2009:13:37:00 +0100]
   "GET /music HTTP/1.0" 200 2324 "http://last.fm/"
   "Mozilla/5.0 (X11; U; Linux i686)"
\end{lstlisting}

@@

\begin{lstlisting}[language=Python,caption={An example map function, parsing
Apache access logs.},captionpos=b]
   def map(key, value):
      log = parse_apache_log(value)
      dimensions = {'useragent' : log.useragent,
                    'path'      : log.path,
                    'status'    : log.statuscode}
      values = {'pageviews' : 1,
                'bytes'     : log.size}

@@

Since the goal is to store projections rather than all combinations of all
dimensions, the data needs to be mangled slightly before sending it down
the pipe to the underlying data store. Dumbo expects a key-value pair from the
mapper. Hadoop will collect all keys and give the reducer a key and an iterator
of values.

@@

Remember, the user map function gives a point in n-space and a list of
measurements. The measurements needs to be extracted and each such measurement
is summed over. Zohmg solves this by wrapping the user map function.

@@

\section{Storage / Data Model}

% TODO: renskriv.
The typical use case for Zohmg is to read one or more column-qualifiers,
which correspond to one or more points in n-space, over a range of time.
Therefor, the keys are made up of the time and unit of the measurement,
and the qualifier queried for represents the point in n-space. More on
this in the data model section.

Explain our data model, the relation between column-families, column-qualifiers
and projections and dimensions.

@@

The data export server is built with a WSGI\footnote{Web Server Gateway
Interface, a HTTP extension where a CGI-like environment is passed around.}
framework, which includes both middleware and application parts. The middleware
applications are responsible for dispatching requests to the corresponding
application or file. WSGI middleware receives a WSGI request and then performs
logic upon this request, before passing the request to a WSGI application.
\cite{paste,definitive_guide_to_pylons} Since the main goal was not to build a
middleware framework from scratch, an existing such was used.

@@

The data export server is built with Python Paste---a middleware
framework---the data server dispatches to the requested application
depending on the URL it was fed. \cite{paste}

@@

\textbf{Replication} &
.

\textbf{Serializing} &
Serializing data is the task of converting an object into a sequence of bits,
making it possible to store on a storage medium or transmitted over a network
link. Rereading the sequence of bits should restore the original state of the
object it was in before the serialization. \\

\textbf{Sharding} &
.
\end{tabular}

@@
